### 2021-03-12 notes/notes.tex
*** empty log message ***

\documentclass{report}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tabularx}
\usepackage[margin=1.5cm]{geometry}
\bibliographystyle{apalike}

@@

\section{The model}
The model is an implementation of \ref{subsec:proposal}. The generalized state is given by:
\[
z_t = \{x, \theta_t, \gamma_t, \omega_t\}
\]
where $x$ is the current position of the hand in Cartesian coordinates, where the origin is the starting point. We assume that motor commands are issued every $\Delta t$ for simplicity. At the beginning of each time interval, the context is inferred combining the cue and the prediction error of the outcome of the previous motor command. We assume further than the cue and the force field are uncorrelated, which yields:
\begin{align}
  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
\end{align}
The terms $p(C_t)$ refer to the prior probability of the context at the beginning of the trial: at the first trial, this refers to prior beliefs over which contexts are more common/likely. At each subsequent trial, it incorporates the belief that has so far been accumulated, given previous trials.

@@

For mathematical tractability (i.e. to avoid circular distributions), we will limit contexts to one where the force points left (angle $\pi$) and one where the force points right (angle 0), in addition to the baseline context. In this scheme, participants need only learn the magnitude of the force for each context. We will assume that for each context, the prior distribution over the magnitude of the force is Gaussian:
\[
p(\gamma) = N(1, 0.1)
\]
We will show that the form of this prior affects learning rate, but not the final learned magnitude.

After the generalized state has been inferred given the information at the beginning of the trial, the next step is to infer the parameters of the context, as in \eref{eqn:update-parameters}:
\begin{equation}
q(C_t | x_t, x_{t-1}, a_{t-1}) \propto p(x_t | x_{t-1}, a_{t-1}, C_t)p(C_t) \label{eqn:context-from-x}
\end{equation}

where $p(C_t)$ is either a prior belief of the likelihood of each context, or the probability inferred up to the previous trial. Note that, because we assumed that the position of the hand is directly observable without any noise, there are no terms for the uncertainty of $x_t$ or $x_{t-1}$.

With this, \eref{eqn:context-from-x} becomes:
\begin{equation}
q(\gamma | x_t, x_{t-1}, a_{t-1}) \propto N(\hat x_t, \sqrt{\Delta t})p(\gamma) \label{eqn:data-dist}
\end{equation}
where $\hat x_t$ is the predicted value of the position $x_t$ given the previous state and action and the force field of the context; see \eref{eqn:dynamics} below. Because we chose the priors $p(\gamma)$ to be Gaussian, the posteriors $q(\gamma)$ after every trial will also be a Gaussian distribution with parameters:
\begin{align}
  \mu_\gamma^{(t)} &= \frac{\mu_\gamma^{(t-1)}\left(\sigma_\gamma^{(d)}\right)^2 + \mu_\gamma^{(d)}\left(\sigma_\gamma^{(t-1)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2} \\
  \sigma_\gamma^{(t)} & = \sqrt{\frac{\left(\sigma_\gamma^{(t-1)}\right)^2\left(\sigma_\gamma^{(d)}\right)^2}{\left(\sigma_\gamma^{(t-1)}\right)^2 + \left(\sigma_\gamma^{(d)}\right)^2}}
\end{align}
where the superscripts $(t)$ and $(t-1)$ refer to the inferred values for the current and previous trials, respectively, and $(d)$ to the parameters in \eref{eqn:data-dist}.

>>>>

\documentclass{report}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tabularx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{todonotes}
\bibliographystyle{apalike}

@@

\section{The model}
The model is an implementation of \ref{subsec:proposal}. The generalized state
is given by:
\begin{equation}
z_t = \{x, \theta_t, \gamma_t, \omega_t\}
\end{equation}
where $x$ is the current position of the hand in Cartesian coordinates,
where the origin is the starting point. We assume that motor commands are
issued every $\Delta t$ for simplicity. At the beginning of each time interval,
the context is inferred combining the cue and the prediction error of the
outcome of the previous motor command. We assume further than the cue and the
force field are uncorrelated, which yields:
\begin{align}
  q(C_t) &= q(C_t | \omega_t)q(C_t | s_t, s_{t-1}, a_{t-1}) \\ \label{eqn:estimated-context}
  q(C_t | \omega_t) &\propto p(\omega_t | C_t)p(C_t) \\
  q(C_t | x_t, x_{t-1}, a_{t-1}) &\propto p(x_t | C_t, x_{t-1}, a_{t-1})p(x_t)p(x_{t-1})p(C_t)
\end{align}
The terms $p(C_t)$ refer to the prior probability of the context at
the beginning of the trial: at the first trial, this refers to prior beliefs
over which contexts are more common/likely. At each subsequent trial, it
incorporates the belief that has so far been accumulated, given previous
trials.

@@

For mathematical tractability (i.e. to avoid circular distributions), we will
limit contexts to one where the force points left (angle $\pi$) and one where
the force points right (angle 0), in addition to the baseline context. In this
scheme, participants need only learn the magnitude of the force for each
context. We will assume that for each context, the agent's belief over the
magnitude of the force is given by a normal distribution:
\begin{equation}
p(\gamma) = N(\mu_f, \sigma_f) \label{eqn:data-dist}
\end{equation}
where $\mu_f$ and $\sigma_f$ are parameters to be estimated at each
trial. Before the experiment begins, the agent will have a prior distribution
over these parameters. A standard Bayesian approach is to choose NormalGamma
priors:
\begin{equation}
p(\mu_f, \sigma_f) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
\end{equation}
where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
model. Because participants are told that there are three contexts (baseline,
left and right), we assume the following hyperparameter values:
\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
  \nu_f^{(0)} &= 0.1 \\
  \alpha_f^{(0)} &= 1 \\
  \beta_f^{(0)} &= 0.5
\end{align}
Note that $\mu_f^{(0)}$ is different for each context and is thus
provided as a vector $(0, -1, 1)$ representing the baseline, left and right
contexts, respectively. For all other hyperparameters, all contexts have the
same value. These values ensure that the prior over the hyperparameters
$(\mu_f, \sigma_f)$ has the mode at $(\mu_f^{(0)}, 1)$. We will show that the
choice of these hyperparameters affects learning rate, but not the final
learned magnitude \todo{Check this}.

The update equations for the magnitud parameters are given by:
\begin{equation}
q(\mu_f, \sigma_f | x_t, x_{t-1}, a_{t-1}) \propto p(x_t |
x_{t-1}, a_{t-1}, C_t)p(\mu_f, \sigma_f) \label{eqn:context-from-x}
\end{equation}
where $p(x_t | x_{t-1}, a_{t-1}, C_t)$ is a Gaussian distribution centered
around $\mu_a$ with standard deviation $\sigma_a$, which is a free parameter of
the model. It is assumed that the standard deviation over outcomes ($\sigma_a$)
is related to that of the parameter $\mu_f$ as follows:
\begin{equation}
\sigma_f = \sigma_a / \nu_f
\end{equation}
which makes $\nu_f$ a parameter to estimate. Because we chose the priors to be
NormalGamma and the likelihood to be Gaussian, the posteriors are also
NormalGamma, and the parameters are updated after one time step via:
\begin{align}
  \mu_f^{(t)} &= \frac{\nu_f^{(t-1)} \mu_f^{(t-1)} + x_t}{\nu_f + 1} \\
  \nu_f^{(t)} &= \nu_f^{(t-1)} + 1 \\
  \alpha_f^{(t)} &= \alpha_f^{(t-1)} + 0.5 \\
  \beta_f^{(t)} &= \beta_f^{(t-1)} + \frac{\nu_f^{(t-1)}}{\nu_f^{(t-1)} +
                  1}\frac{\left(x - \mu_f^{(t-1)}\right)^2}{2}
\end{align}
